The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-01 paper.tex
Typos, formatting in introduction

\def\journal{Cognitive Science}%%% write here for which journal %%%
\def\DOI{}
\def\articleType{Research Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Fink {et~al.}} %use et al only if is more than 1 author
\def\Authors{Julia Fink\,$^{1*}$, S\'{e}verin Lemaignan$^{1}$, Claire Braboszcz$^{2}$ and Pierre Dillenbourg$^{1}$}
\def\Address{$^{1}$Computer-Human Interaction in Learning and Instruction (CHILI) \\ Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
\vspace{0.25cm}
$^{2}$Laboratory for Neurology and Imaging of Cognition (LabNIC) \\
Universit\'{e} de Gen\`{e}ve \\ CH-1211 Gen\`{e}ve, Switzerland 
}
\def\corrAuthor{Julia Fink \vspace{1mm}}
\def\corrAddress{EPFL - CHILI, RLC D1 740, Station 20, CH-1015 Lausanne, Switzerland}
\def\corrEmail{julia.fink@epfl.ch}

@@

A phenomenon that is often studied related to the human-like design of 
robots is anthropomorphism. While people's tendency to perceive human-like 
characteristics in robots is a commonly discussed trait of 
human-robot interaction (HRI), the phenomenon paradoxically lacks formal 
grounds. Based on findings from several empirical studies and a 
comprehensive synthesis of literature, this article attempts at extending 
the current understanding of anthropomorphism. In accordance with theories 
from social psychology, cognitive science, and neuroscience, we suggest to 
formalize qualitative differences and dynamics over time in how 
humans anthropomorphize non-human agents. Based on several key factors, 
namely the 
design of the non-human agent, the individual characteristics of the human 
user, and the context of use, we found that anthropomorphism changes in a 
non-linear way with growing experience of the user with the system. To 
account for these variances, we propose an initial model that we call 
\textit{``Dynamics of Anthropomorphism''}. Focusing on human-robot 
interactions, we present a cognitive interpretation that considers 
the described dynamics. Finally, we would like to discuss how others can 
apply our model to measure and describe better the complexity in human's 
tendency to perceive human-like characteristics in non-human agents.

@@

The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like
designs of objects and non-human agents can evoke both positive and negative feelings 
of affinity in the human observer / user. A positive valence reflects the experience of 
emotional engagement with and feelings of empathy for the non-human agent
\citep{cheetham_human_2011}. This in turn may lead to an effective (human-like and 
human-social) interaction with the non-human agent, and increase people's acceptance of
it. 
However, at a certain point as the non-human agent
approaches but fails to closely resemble a human, the evoked valence in the
human user suddenly turns into a negative uncanny feeling of what
\cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn,
might lead to rejection of the non-human agent. The Uncanny
Valley hypothesis predicts that with further increasing the human-likeness of
the non-human agent such that it appears so realistic that it is not
distinguishable from a human but perceived to be human, the effect is again a
positive one, reflecting feelings and engagement that are similar as those
evoked by a human-human interaction. 
Of course, developers and designers of artificial agents
(\eg virtual characters, robots) try to exploit the positive effects of 
\textbf{anthropomorphic form} and avoid ``falling'' into this Uncanny Valley.
Consequently, there is a considerable research effort to
better understand the boundaries of anthropomorphic design and their effects
on the interaction and experience. To date, findings have been however
inconsistent \citep{cheetham_human_2011}.

A phenomenon that is often studied related to the human-like design of objects and 
artificial agents is \textbf{anthropomorphism}.
Along this line, this article aims at fostering a better understanding of
anthropomorphism (noted \Ant in this paper) in human-robot interaction 
by considering the phenomenon
as a whole. Robotics researchers often tend to believe that
anthropomorphism only describes a static set of human-like features of a robot
(like shape, speech capabilities, facial expressions, etc.). We refer to these
characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}, and propose that \emph{anthropomorphism} in 
general refers to a \emph{social phenomenon} that emerges from the (real or
imagined) \emph{interaction} between a non-human agent (\eg a robot) and a 
human~\citep{persson_anthropomorphism_2000}. According to~\cite{epley_when_2008},
anthropomorphism includes that the human perceives, for instance, emotional states, motivations, or 
intentions in the non-human agent, and tends to ascribe those qualities to it.
As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with artificial agents such as robots.

@@

To support the reflection on variances and dimensions of anthropomorphism, it may be 
useful to consider a few experimental
results. We studied anthropomorphism in several short- and long-term
human-robot interaction studies that we carried out with embodied robots, both
in ecologically valid real-world environments and controlled lab settings. Those
experiments are presented in details in previously published
articles~\citep{fink_anthropomorphic_2012, fink_living_2013, fink2014which,
lemaignan2014dynamics}. We only highlight here a few qualitative results that
underline the complexity of the phenomenon.

\begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-activities.png}
                \caption{Activities related to Roomba, n = 9 households}
                \label{fig:roomba-activities}
        \end{subfigure}%
        \hspace{0.5cm} 
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-perception.pdf}
                \caption{People's perception of Roomba, n = 15 participants}
                \label{fig:roomba-perception}
        \end{subfigure}
    \caption{Usage and perception of the Roomba vacuum-cleaning robot over a 6 months period. Qualitative data suggest that people's activities related to the robot as well as their perception of the robot change over time.}
    \label{fig:roomba}
\end{figure}

Figures~\ref{fig:roomba-activities} and~\ref{fig:roomba-perception} give a first
perspective on how naive users interact and perceive robots over a long period
of time. Here, \emph{Roomba} vacuum cleaning robots were given to nine
households for a 6-months period. The aim was to explore how people use and
perceive the robot over time, and how they generally live with the Roomba in
their home. At various time points, users were asked to rate their perception of
different features of the robot on a 7-point rating scale. As
Figure~\ref{fig:roomba-perception} shows, some features remained stable over
time (\emph{ease of use} and \emph{fun}), and others decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one does not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on one's life (and the level of activity pictured in
Figure~\ref{fig:roomba-activities} qualitatively confirms this intuition). Even
if one can argue that the Roomba is not designed to foster interaction, it
raises a first question: how to sustain engagement between a robot and a human
over a long period of time? And, as a pre-requisite to answer, do we actually
understand the psychological phenomenon, and can we effectively account for the
dynamics of these interactions?

Those questions are certainly not new, but paradoxically, while anthropomorphism
is a commonly discussed trait of human-robot interaction, it appears
that we lack formal grounds and models that would account for the long-term
dynamics of these interactions.

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.75\textwidth]{ranger}
                \caption{Interaction scenario of the \emph{Ranger} domino study: the robot toy box Ranger transports domino tiles between the two children}
                \label{fig:ranger-expe}
        \end{subfigure}%
        \hspace{0.5cm}
        \begin{subfigure}[b]{0.48\columnwidth}
            \includegraphics[width=\textwidth]{ranger-interactions-vs-qualitative-score}
            \caption{We found a significant negative correlation ($p=0.048$) between the
    number of interactions and the qualitative anthropomorphic projections. n = 13 groups}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Children`s interaction with Ranger, and their anthropomorphic projections onto the robot. Overall, large variations of both children`s behavior and their perception of the robot were found but seemed to be correlated to each other.}
    \label{fig:ranger}
\end{figure}

%\fixme{It was written as caption of fig:ranger: ``Level of interaction with a semi-anthropomorphic robot is \emph{not} a predictor of anthropomorphic projections.'' Comment JF: this statement does not make sense to me as we found a significant negative correlation between the two.}

@@

These first remarks lead to the main question we attempt to address in this
article: how can we understand \emph{anthropomorphism} as a whole, \hl{and in such a
way that is support improved human-robot interactions?}\fixme{JF: something's wrong with this sentence}

We propose to discuss anthropomorphism from three
complementary perspectives:

%Based on an extensive literature review~\citep{fink_anthropomorphism_2012},
%several field and lab studies previously mentioned, and insights from the
%cognitive neuroscience, we propose to discuss anthropomorphism from three
%complementary perspectives.

@@

%\paragraph{Two hypotheses}
Our research question also translates to two
hypotheses, which structure the reflections of this article:

\begin{enumerate} 
\item ~A person's tendency to anthropomorphize a robot
    is based on several factors, including (but not limited to) (1) the
    design of the robot, (2) the personal characteristics of the human user,
    and (3) the context and situation of use. For instance, a social context
    and purpose of the robot increase anthropomorphism: \textit{anthropomorphism
    as a multi-factor phenomenon}.

\item ~A person's tendency to anthropomorphize a robot
    evolves in a non-linear way over time and with growing experience that the
    person has with the robot. For instance, a more experienced user will
    perceive the robot as more familiar and predictable, and in turn
    anthropomorphizes the robot less: \textit{anthropomorphism as a dynamic
    phenomenon}.

@@

%\begin{inparaenum}[\itshape 1)\upshape]

The remainder of this article is organized as follows.
According to the three perspectives that we take, 
we first (in Section
%\ref{sec:overview-studies}, we give a brief overview on the various studies that
%we conducted to investigate anthropomorphism, namely \item the Roomba study,
%\item the Forum study, \item the Ranger domino study. Next, Section
\ref{sec:anthropomorphism}) aim to establish a common ground in the
understanding of anthropomorphism. We review literature from several relevant domains and integrate the different explanations into an understanding of
anthropomorphism as a multi-factor phenomenon.
Based on this, Section~\ref{sec:our-ideas} introduces our initial model of the
\textit{Dynamics of Anthropomorphism}. This model describes the variances
and dynamics of people's tendency to anthropomorphize an artificial agent,
such as a robot. 
Then Section \ref{sec:cognition-neuroscience} presents a
cognitive and neuroscience perspective on anthropomorphism to consolidate
our considerations and the proposed model.
This is followed by a discussion
about the limitations of our approach (Section \ref{sec:discussion}) and the
existing and alternative ways to measure anthropomorphism in human-robot
interactions.
Finally, Section \ref{sec:conclusion} concludes this article
and offers starting points for further research on anthropomorphism and the
effect of the human-like design approach in robotics.

%\end{inparaenum}

@@

\emph{Familiarization}, in this context, means \emph{to get acquainted with}.
It is not to be confused with the \emph{familiarity thesis} that relates to the
projection of already known cognitive models onto the robot. We discuss this other
aspect at section~\ref{sec:cognitivemodel}.

@@

Common observation of naive people (children or adults) interacting with robots
shows that unexpected behaviors of the robot can have a notable impact on
interaction. \hl{cite literature, e.g. cheating robot, disagreeing, disobeying,
HRI2013 paper}
This is supported by the results from \citet{Waitz2010} seen in section
\ref{sec:explanations} that people attribute more easily anthropomorphic
features to artefacts when they have unpredictable behaviors.

>>>>

\def\journal{Cognitive Science}%%% write here for which journal %%%
\def\DOI{}
\def\articleType{Research Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Fink {et~al.}} %use et al only if is more than 1 author
\def\Authors{Julia Fink\,$^{1*}$, S\'{e}verin Lemaignan$^{1}$, Claire Braboszcz$^{2}$ and Pierre Dillenbourg$^{1}$}
\def\Address{$^{1}$Computer-Human Interaction in Learning and Instruction
    (CHILI) \\ \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
\vspace{0.25cm}
$^{2}$Laboratory for Neurology and Imaging of Cognition (LabNIC) \\
Universit\'{e} de Gen\`{e}ve \\ CH-1211 Gen\`{e}ve, Switzerland 
}
\def\corrAuthor{Julia Fink \vspace{1mm}}
\def\corrAddress{EPFL - CHILI, RLC D1 740, Station 20, CH-1015 Lausanne, Switzerland}
\def\corrEmail{julia.fink@epfl.ch}

@@

Anthropomorphism is a phenomenon that is often studied related to the human-like
design of robots. While people's tendency to perceive human-like 
characteristics in robots is a commonly discussed trait of 
human-robot interaction (HRI), the phenomenon paradoxically lacks formal 
grounds. Based on findings from several empirical studies and a 
comprehensive synthesis of literature, this article attempts at extending 
the current understanding of anthropomorphism. In accordance with theories 
from social psychology, cognitive science, and neuroscience, we suggest to 
formalize qualitative differences and dynamics over time in how 
humans anthropomorphize non-human agents. Based on several key factors, 
namely the 
design of the non-human agent, the individual characteristics of the human 
user, and the context of use, we found that anthropomorphism changes in a 
non-linear way with growing experience of the user with the system. To 
account for these variances, we propose an initial model that we call 
\textit{``Dynamics of Anthropomorphism''}. Focusing on human-robot 
interactions, we present a cognitive interpretation that considers 
the described dynamics. Finally, we would like to discuss how others can 
apply our model to measure and describe better the complexity in human's 
tendency to perceive human-like characteristics in non-human agents.

@@

The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like
designs of objects and non-human agents can evoke both positive and negative
feelings of affinity in the human observer / user. A positive valence reflects
the experience of emotional engagement with and feelings of empathy for the
non-human agent \citep{cheetham_human_2011}. This in turn may lead to an
effective (human-like and human-social) interaction with the non-human agent,
and increase people's acceptance of it.  However, at a certain point, as the
non-human agent approaches but fails to closely resemble a human, the evoked
valence in the human user suddenly turns into a negative uncanny feeling of what
\cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn,
might lead to rejection of the non-human agent. The Uncanny Valley hypothesis
predicts that with further increasing of the human-likeness such as it appears
so realistic that it is not distinguishable from a human and actually perceived
to be human, the effect is again a positive one, reflecting feelings and
engagement that are similar as those evoked by a human-human interaction.
Developers and designers of artificial agents (\eg virtual characters, robots)
try to exploit the positive effects of \textbf{anthropomorphic form} and avoid
``falling'' into this Uncanny Valley, and consequently, there is a considerable
research effort to better understand the boundaries of anthropomorphic design
and their effects on the interaction and experience. To date, findings have
remained however inconsistent \citep{cheetham_human_2011}.

A phenomenon that is often studied related to the human-like design of objects
and artificial agents is \textbf{anthropomorphism}.  Along this line, this
article aims at fostering a better understanding of anthropomorphism (noted \Ant
in this paper) in human-robot interaction by considering the phenomenon as a
whole. Robotics researchers often tend to believe that anthropomorphism only
describes a static set of human-like features of a robot (like shape, speech
capabilities or facial expressions). We refer to these characteristics as the
\emph{anthropomorphic design} of the robot~\citep{fink_anthropomorphism_2012},
and propose that \emph{anthropomorphism} in general refers to a \emph{social
phenomenon} that emerges from the (real or imagined) \emph{interaction} between
a non-human agent (\eg a robot) and a
human~\citep{persson_anthropomorphism_2000}. According
to~\cite{epley_when_2008}, anthropomorphism includes that the human perceives,
for instance, emotional states, motivations, or intentions in the non-human
agent, and tends to ascribe those qualities to it.  As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with artificial agents such as robots.

@@

To support the reflection on variances and dimensions of anthropomorphism, it
may be useful to first consider a few experimental results. We studied
anthropomorphism in several short- and long-term human-robot interaction studies
that we carried out with embodied robots, both in ecologically valid real-world
environments and controlled lab settings. Those experiments are presented in
details in previously published articles~\citep{fink_anthropomorphic_2012,
fink_living_2013, fink2014which, lemaignan2014dynamics}. We only highlight
here a few qualitative results that underline the complexity of the
phenomenon.

\begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-activities.png}
                \caption{Activities related to Roomba, $n = 9$ households}
                \label{fig:roomba-activities}
        \end{subfigure}%
        \hspace{0.5cm} 
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-perception.pdf}
                \caption{People's perception of Roomba, $n = 15$ participants}
                \label{fig:roomba-perception}
        \end{subfigure}
    \caption{Usage and perception of the Roomba vacuum-cleaning robot over a 6
    months period. Qualitative data suggest that people's activities related to
    the robot as well as their perception of the robot change over time.}

    \label{fig:roomba}
\end{figure}

Figures~\ref{fig:roomba-activities} and~\ref{fig:roomba-perception} give a first
perspective on how naive users interact and perceive robots over a long period
of time. Here, \emph{Roomba} vacuum cleaning robots were given to nine
households for a 6-months period. The aim was to explore how people use and
perceive the robot over time, and how they generally live with the Roomba in
their home. At various points in time, users were asked to rate their perception
of different features of the robot on a 7-point rating scale. As
Figure~\ref{fig:roomba-perception} shows, some features remained stable over
time (\emph{ease of use} and \emph{fun}), and others decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one does not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on one's life (and the level of activity pictured in
Figure~\ref{fig:roomba-activities} qualitatively confirms this intuition). Even
if one can argue that the Roomba is not designed to foster interaction, it
raises a first question: how to sustain engagement between a robot and a human
over a long period of time? And, as a pre-requisite to answer, do we actually
understand the underlying psychological phenomenons that explain the decrease of
interest, and can we effectively account for the dynamics of these phenomenons?

Those questions are certainly not new, but paradoxically, while anthropomorphism
is a commonly discussed trait of human-robot interaction, it appears that we
still lack formal grounds and models that would account for the long-term
dynamics of the interaction.

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.75\textwidth]{ranger}
                \caption{Interaction scenario of the \emph{Ranger} domino study: the robot toy box Ranger transports domino tiles between the two children}
                \label{fig:ranger-expe}
        \end{subfigure}%
        \hspace{0.5cm}
        \begin{subfigure}[b]{0.48\columnwidth}
            \includegraphics[width=\textwidth]{ranger-interactions-vs-qualitative-score}
            \caption{We found a significant negative correlation ($p=0.048$) between the
    number of interactions and the qualitative anthropomorphic projections. $n = 13$ groups}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Children's interaction with Ranger, and their anthropomorphic
        projections onto the robot. Overall, large variations of both children's
        behavior and their perception of the robot were found but were
        correlated to each other.}

    \label{fig:ranger}
\end{figure}

@@

These first remarks lead to the main question we attempt to address in this
article: how can we understand \emph{anthropomorphism} as a whole, and in such a
way that it supports improved design of human-robot interactions?

To this end, we propose to discuss anthropomorphism from three complementary
perspectives:

@@

Our research question also translates to two
hypotheses, which structure the reflections of this article:

\begin{enumerate} 
\item ~A person's tendency to anthropomorphize a robot
    is based on several factors, including (but not limited to) (1) the
    design of the robot, (2) the personal characteristics of the human user,
    and (3) the context and situation of use. For instance, a social context
    and purpose of the robot increase anthropomorphism: \textit{anthropomorphism
    is a multi-factor phenomenon}.

\item ~A person's tendency to anthropomorphize a robot
    evolves in a non-linear way over time and with growing experience that the
    person has with the robot. For instance, a more experienced user will
    perceive the robot as more familiar and predictable, and in turn
    anthropomorphizes the robot less: \textit{anthropomorphism is a dynamic
    phenomenon}.

@@

The remainder of this article is organized as follows.  According to the three
perspectives that we take, we first (in Section \ref{sec:anthropomorphism}) aim
to establish a common ground in the understanding of anthropomorphism. We review
literature from several relevant domains and integrate the different
explanations into an understanding of anthropomorphism as a multi-factor
phenomenon.  Based on this, Section~\ref{sec:our-ideas} introduces our initial
model of the \textit{Dynamics of Anthropomorphism}. This model describes the
variances and dynamics of people's tendency to anthropomorphize an artificial
agent, such as a robot.  Then Section \ref{sec:cognition-neuroscience} presents
a cognitive and neuroscience perspective on anthropomorphism to consolidate our
considerations and the proposed model.  This is followed (Section
\ref{sec:discussion}) by a discussion about the limitations of our approach and
the existing and alternative ways to measure anthropomorphism in human-robot
interactions.  Finally, Section \ref{sec:conclusion} concludes this article and
offers starting points for further research on anthropomorphism and the effect
of the human-like design approach in robotics.

@@

\emph{Familiarization}, in this context, means \emph{to get acquainted with}.
It is not to be confused with the \emph{familiarity thesis} that relates to the
projection of already known cognitive models onto the robot. We discuss this other
aspect at section~\ref{sec:cognition-neuroscience}.

@@

Common observation of naive people (children or adults) interacting with robots
shows that unexpected behaviors of the robot can have a notable impact on
interaction. \hl{cite literature, e.g. cheating robot, disagreeing, disobeying,
HRI2013 paper}
This is supported by the results from \citet{Waytz2010} seen in section
\ref{sec:explanations} that people attribute more easily anthropomorphic
features to artefacts when they have unpredictable behaviors.
